Pattern mining in spatio-temporal datasets is a really relevant subject in the academia and the industry nowadays, due
to its wide applicability in helping to solve real-world problems. Many of them can be found in the context of Smart
Cities, like Traffic Management, Surveillance and Security and City Planning, to name a few. Among the various
spatio-temporal patterns that one can extract from a spatio-temporal dataset, the flock pattern is one that has gained a
lot of attention, because of its intrinsic relation with the aforementioned problems. A lot of work has been done in the
academia, in order to provide algorithms able to identify the flock pattern. However, none of them could perform that
task efficiently nor be able to scale well when a large dataset was the analysis target. Additionally, we found that
there was no system architecture proposal that could be simple and modular enough to be used in that spatio-temporal
pattern detection problem. Given that context, this dissertation proposes a modular system archicture that can be used
to solve flock pattern mining problems and possibly be reused to other spatio-temporal mining experiments. We then use
such architecture as the infrastructure to implement an efficient flock detection algorithm, aiming at achieving
considerable gains in execution time without compromising accuracy, thus targeting real-time deployment and online
processing in Smart Cities. Last, but not least, we remodel our algorithm in order to take advantage of multi-core
architectures present in modern computers. Our results indicate that our proposal outperforms the current
state-of-the-art techniques, by achieving 99\% CPU time improvement. Moreover, with our multi-thread model, we were able
to reduce the processing time of our proposed algorithm by 96\% in some cases. We prove the efficiency of our solution
by performing evaluation with both real and synthetic large datasets.

\begin{keywords}
Smart Cities, Pattern Mining, Spatio-temporal data, Flock Pattern
\end{keywords}
